Stack Safety as a Security Property

### 2020-09-09 rebuttal.txt

We thank the reviewers and in particular reviewer C for their thoughtful
comments.  We first discuss the most important points raised in
reviews. Then, in an optional appendix, we respond point-by-point to each
individual piece of reviewer feedback.

MOTIVATION

Reviewer C is particularly concerned with the way this work is
motivated. Our original goal in this work wasn't to formalize a stack safety
notion that is specific to micropolicies, but in hindsight we agree that the
result may appear to be tailored to that particular enforcement mechanism,
though we believe this is more a matter of presentation choices and
exposition-helping simplifications than a fundamental limitation. In
particular, we believe that the enforcement model can be used (with some
effort) to describe capability-based enforcement mechanisms, as we state in
Section 2.  However, given that this effort didn't make it into the final
product, we agree that the paper as it stands would be more accurately
described as an abstract property capturing micropolicy enforcement rather
than the all-encompassing stack safety notion we set out to state, and
intend to do so for a final version.

BCP: I.e., we plan to limit our claims rather than fixing the exposition?
Why not promise the latter (and lay out what we would do)?
LEO: Going with "we'll limit our claims and add the related work comparison"
seems more doable (and fair) in a revision round than promising to do
new work (or work that we started doing and then it was clear we couldn't
work for the deadline).

@@

RELATED WORK

All reviewers also wanted additional discussion of related work.  We agree
that there should be a clearer comparison to both Skorstengaard papers,
using the informal criteria proposed there. We plan to include a discussion
like the following:

(1) On intuition: The LSE part of the property we present aims to
characterize stack safety in security terms. For readers familiar with
integrity/confidentiality/noninterference, the eager version should be
fairly intuitive. The lazy/observable properties, are less intuitive,
but more interesting in that they precisely characterize violations to
the more intuitive view of stack safety that can still be considered
safe to an external observer. The WBCF part simply formalizes standard
behavior of stack-based control flow.

@@

By contrast, Skorstengaard relies on different prior intuitions:
for readers familiar with the behavior of pure stacks (or rather the
built-in oLCM stack), their property is more intuitive. For readers
versed in security, we believe there is value in a different
perspective.

(2) useful for reasoning: In order to reason about actual programs we
would indeed need to incorporate a "Meaning of Memory Safety"-style
frame rule. Our goal with this paper was to distill the essence of
stack safety without adding the complex (but solved) overhead of
dealing with heap locations. But the properties (and the testing
framework behind it) were designed with [Azevedo et al] in mind.

(3) reusable in compiler chains: As you point out, this is not really
applicable, since this is a criterion tailored to what Skorstengaard
is aiming to do. We might consider what the equivalently tailored
criterion would be for us -- usable in "dynamic enforcement
environments" with potentially unsafe commodity compilers.

APT: Reviewer C interprets (4) to mean:
"This criterion asks for an argument that any property of a program
that relies on stack safety can be proven with the characterization of stack safety."
Since we're in the business of _defining_ stack safety, this seems pretty circular to me!
Maybe the point is that we should not omit any characteristic of
everyone's informal notion of stack safety (for the set of language features
we support -- which is the business of (5)).  Can we claim that (informally, of course)?

(4) complete: This criterion seems a bit circular. It asks for being
able to prove any property of a program that relies on stack safety
with a given characterization of stack safety - but we're in the
buisness of _defining_ stack safety in the first place. If the point
is that our definition should capture characteristics of established
informal notions of stack safety (subject to the limitations of the
language features our formalization supports - see 5), our properties
should apply to any such reasonable program.

@@

> - The proposed notion of stack safety is too strong to be enforced by existing stack protection mechanisms.
>
>   [Roessler and DeHon 2018]'s "lazy tagging and clearing" is itself not stack-safe according to this paper.
>   To understand the mechanism, this paper additionally defines the "observable" versions of stack safety.
>   But then why you didn't use the observable version as the authoritative definition?

We propose two definitions of stack safety: one that is easier to understand
because it corresponds to a more naive view of what stack safety intuitively
means, and a slightly more complex and permissive one that (as Roessler and
Dehon showed) can be implemented more efficiently.  We believe both are of
interest.

> - This paper is lacking proper discussion on related work.
>
>    On the contrary to this paper's observation (L25), there're a few formalizations of stack safety (and rely/guarantee conditions for function invocation).
>       For example, see CompCert's assumption on function calls: https://github.com/AbsInt/CompCert/blob/94558ecb3e48261f12c644045d40c7d0784415e0/common/Events.v#L619

BCP: Obviously, this has to be addressed strongly and concretely.
LEO: I'm not really following the CompCert definition. Has anybody taking a crack at this?

@@

>   Also, they should discuss language features (more than the ones currently in section 9, like closures and dynamic code generation and discuss how their approach could potentially scale to it in the long term.
> * discuss in more detail the fundamental limitations of their approach, particularly the assumption of a static code layout.
> * provide a clear, maximally objective/neutral statement of the benefits of their approach w.r.t. Skorstengaard et al. 2019b

We agree that there should be a much more thorough comparison with Sk2019a/b. See the summary above for more details on what we plan to include.

> The authors do not discuss any attacker model, both for their properties and for their enforcement technique.
> Who can modify the micropolicy tables?
> Do you have to trust the operating system or just the hardware?
> Is the attacker capable of running arbitrary assembly code (like in Skorstengaard et al, 2019ab) or just inject stack modifications like in Abadi et al., 2005?

See ATTACKER MODEL in the summary above.

> Another odd aspect of this paper is the fact that it doesn't propose one but three increasingly strong properties that characterize stack safety (see l86-96).
> What does this mean: why do we want stack safety to mean different things and which meaning is appropriate in which setting?

We view these stack safety properties as a spectrum. The main property
strives to capture an intuitive notion of stack safety, by relying on
abstract notions of observable behavior.  The stronger, lockstep
enforces a more conservative notion of stack safety, where almost any
digression from a canonical execution is considered unsafe. On the
other hand, the weaker lazy variant pushes the limit of the standard
intuition by allowing executions that would otherwise be unsafe, as
long as their effects are not actually observable. We think all
three are interesting in their own right and reveal interesting trade
offs in terms of enforcement or reasoning efficiency.

> Do the three variants score equally well on the above criteria?
> - For the "lockstep" variant, I get the impression that this is a proof-technical definition that is only used for proving the regularly stack-safety property for the micropolicy enforcement and not really important/useful in itself: is this also how you see it?

The lockstep variant is a slightly more technical artifact that is (1)
easy to enforce and (2) easy to test, but that is interesting in its
own right because it captures a very conservative notion of stack
safety---while still allowing the majority of reasonable stack-safe
programs.

> - The lazy variant of stack safety strikes me as problematic because of several reasons:
>   * It seems very specific to exactly one enforcement technique.
>     There isn't a single argument in favour of this property except that it is enforced by this one enforcement technique.
>   * The property is quite complex to formalize and understand. I found it hard to motivate myself to study it in detail, because it is rather complicated and so specific to this one type of enforcement.
>   * The lazy enforcement seems to create new side-channels in the presence of an attacker being invoked by an adversary.
>     Can't the lazy enforcement be exploited by such an adversary to cause the trusted code to halt in the middle of its execution, possibly in an inconsistent state?
>     This reminds me of page-table based attacks in the context of SGX, see, for example, Xu et al. (https://ieeexplore.ieee.org/abstract/document/7163052?casa_token=0ycUzVYD0FoAAAAA:bi8eLFwzqvCnRjtcAWD-y9xG473kJeLQdZaLWEbU43cpAwvGldBu840sXWtcC2mnOnOqGsUPruc).

Similarly, the lazy variant is interesting because it is a property
that is as liberal as possible in the behaviors it allows. Compared to
the other properties, it's much harder to reason about or test (as we
have to consider the entire execution trace), but it's simpler to enforce
and more permissive.

> I'm also a bit unconvinced about the paper's motivation.
> Based on comments throughout the paper, they seem to be looking for a property that describes in detail how the stack looks in memory and how the different memory locations are allowed to be modified or how it is allowed to influence the running program.
> I get the impression they are motivated to find such a property in order to be able to derive a policy that can be enforced on a micropolicy machine.
> In my opinion, this is not an unreasonable goal, but it is a different goal than finding an abstract definition of stack safety that is intuitive, abstract, and maximally applicable to diverse enforcement mechanisms and complicated features like dynamic code generation, varargs etc.
> Properties from related work (Skorstengaard et al., 2019b) seem to strive for a more abstract statement along the lines of "the program should behave the same as when it were running on a machine with a native stack, even in the presence of an adversary" and avoids going into the matter of precisely which memory cell is allowed to be modified at which moment.
> I have the impression the paper would be improved by being more clear about the real goal of the work: finding a more concrete specification of stack safety that is closer to implementable micro-policies.
> To this end, the authors should be open about this and remove statements that suggest their goal is different, for example l29-31 "there is still room to craft a more abstract characterization of stack safety that is independent of capabilities or any other specific enforcement technique".

See MOTIVATION in the summary above.

> General comments:
> * There is some important missing related work on formalizing stack safety:
>   - Martín Abadi, Mihai Budiu, Úlfar Erlingsson, and Jay Ligatti. 2005. A Theory of Secure Control Flow. In Formal Methods and Software Engineering. Springer Berlin Heidelberg.
>   - Martín Abadi, Mihai Budiu, Úlfar Erlingsson, and Jay Ligatti. 2005. Control-flow Integrity. In Conference on Computer and Communications Security. ACM. https://doi.org/10.1145/1102120.1102165
>   This probably will not compromise novelty of the paper's contributions, but I think it must be cited and compared against.
>
> * l17-21: strange that you don't mention CFI here or on l149-150 where you mention "[stack safety enforcement using] code rewriting".
>
> * l23: "defined in a negative way": it's quite unclear what you mean exactly with this phrase.
>   For example, does Abadi2005 (see above) fall under this phrasing?
>   Skorstengaard 2019b can also be described negatively as "Full abstraction should not be broken." and the same applies to this paper.
>   Can you more precisely describe what you mean?
>
> * l197-244: these figures are very clear and helpful, nice...
>   perhaps you could additionally annotate l203-211 as "main", l219-225 as "f" and l233-237 as "g", to clearly point out the link to Fig. 1
>
> * l299-300 and Fig.2: I find it confusing that you use JALR (jump-and-link-register) when you really mean JR (jump to register).
>   IIUC, this is because RISC-V doesn't actually offer JR because it can be emulated as JALR with r0 as the target for storing the return pointer.
>   However, I would prefer if you simply pretend that a direct JR instruction exists.
>
> * l318: "components K": odd name. why not locations as in compiler terminology?
>
> * l426: "and a machine that steps for a long time before producing some output": actually, it's even more strange: the observer cannot distinguish between a machine that steps forever without producing output and a machine that *immediately* produces output
>
> * l483: err.. shouldn't you also require that the stacks below SP are the same in m_c and m_r and that there have been no returns beyond SP between m_c and m_r?
>
> * l515-516: here, I found myself confused because I was interpreting your policy as an information flow policy, so I thought you were allowing to write data from uninitialized stack locations to out.
>
> * l563: "Intuitively, the caller expects to find their private data untouched after the callee returns": what about settings where closures can access the stack frame they were defined in?
>
> * l688 "it also holds for the toplevel execution trace to account for unallocated memory": I don't understand this, please elaborate.
>
> * l1110-1126: I don't understand what you're saying here: Roessler and Dehon's lazy tag policies proposals do not actually give you lazy stack safety?
>   Is there one enforcement technique that satisfies this lazy version of stack safety?
>
> Minor comments:
>
> * l153: "they been applied"
>
> * l298: "[jal] atomically simultaneously the return address in a register": verb missing
>
> * l368: "coinductively": I would say "corecursively" here
>
> * l393: "T in Traces(A) x Obs": do you mean "Traces(T x Obs)"?
>
> * l405: "if f be a predicate": if f is a predicate?
>
> * l537: "MP-trace. say MP": capitalize Say?
>
> * l559: "marked HC in C": HI?
>
> * l687: "TraceConf justret m_c MP_call M_call": parentheses around justret m_c
>
> * l872: "with regard to the original machine state": what does this mean?
>
> Questions for the response period
> ---------------------------------
> * You seem to say/suggest that the approach of Skorstengaard 2019b to defining stack safety is specific to capability machines.
>   Do you really think so and if so, why do you think their approach could not be applied to a different enforcement mechanism?

### 2020-09-09 rebuttal.txt

Sk2019b _define_ stack safety via a capability machine semantics that is supposed to embody stack
safety "by construction," and relating it to a standard semantics for the same machine via a full abstraction result.
We did not mean to imply that the same idea could not work for other machine models or enforcement mechanisms (although
this generalization is not actually in Sk2019b, even as suggested future work). Rather, we wished to distinguish our
approach from theirs by emphasizing that we define stack safety based not on any "correct by construction"
machine tied to some flavor of enforcement,  but rather on more abstract notions of observable behavior.

> * What exactly is the benefit of your approach to defining stack safety w.r.t. the approach of Skorstengaard 2019b?
>   Do you think it can be applied to more or less languages?
>   Do you think it is better suited or less well suited for reasoning about programs which depend on stack safety?
>   Do you think it is etc.

See the full comparison in terms of the criteria in the summary.

> * Do you think you could show safety of a specific example program that depends on stack safety (like the awkward example) and if so, how would such a proof work?

LEO: This is the (4) criterion I guess again. I'm not sure exactly what this means.

> * Do you think your approach to defining stack safety could be extended to languages with the following features (and if so, how?):
>   - varargs?  (seems problematic for call map)

Yes. The call map serves two purposes: identifying call points and
allowing for an easy calculation of the stack pointer at the time of
the call. We could keep the call map essentially as a set and use a
frame pointer (or a similar abstraction) for the latter functionality.
Keeping the number of arguments in the map is certainly convenient,
but not necessary.

>   - dynamic code generation (seems problematic for the function map, call map and return map)
>   - dynamic linking (seems problematic for the function map, call map and return map)
>   - self-editing code (seems problematic for call map and return map)

LEO: Polish needed here.
These three extensions would indeed be challenging. We could
potentially handle generated code by using special contours for the
instructions that generate it, which would correspond to a map that
allows some instructions to update the maps, including potentially
the map itself.

>   - tail calls (seems potentially okay, but I'd like to be sure)

Yes. The only minor problem with tail calls is characterizing the
return - once the point of context switch is determined, the statement
of stack safety remains largely the sam.

>   - exceptions and stack unwinding

LEO: Hmmm. I don't remember ever discussing this.

>   - closures (seems potentially problematic for precision of the entry map)

See below.

>   How would you extend your approach to define stack safety in the presence of these features?
>
> * Do you think your approach implies the following properties?
>   - temporal safety in the presence of closures (i.e. closures should only be invokable by the adversary after the adversary first receives a reference to them)
>   - encapsulation of closures: it should not be possible for the adversary to invoke a closure with a different environment than the one they received?
>   If not, do you think it should imply these properties and if so, how do you think it could be extended?

Adding closures would require: 
(1) identifying code points where closures are created
(2) adding machinery to track association of closure identities, entry points, and components
    containing valid closure pointers (including pointer provenance, as with stack-derived pointers)
(3) extending WBCF to require that closure invocations respect the closure's designated entry point
    and use a valid closure pointer (ensuring temporal safety and encapsulation)
(4) extending contours to treat some components as sealed, associated with closure identities
(5) on closure invocation, set contour for encapsulated components to LC/LI, other components as normal for stack safety
    (allowing access to closure)
(6) otherwise treat sealed components as HC/HI except that from a caller's perspective, they might change during
    nested calls (monotonicity is lost)

So to sum up, the answers are yes, we believe that to add closures,
natural extensions to WBCF and LSE would imply the properties
mentioned.

@@

> * Are the lazy and strong variant of your properties specifically intended for the corresponding micropolicy-based enforcement techniques or do you believe they are more broadly applicable?

As discussed in the summary above and in your detailed comments, we
did not intend for the properties to be micropolicy specific, but a
spectrum of stack safety properties presenting an interesting
trade-off in terms of enforcement or reasoning efficiency. But we
agree that the current exposition does not convey what we set out to
do, and that the overall motivation of the included work would be
better portrayed as a formal characterization of micropolicy-enforced
stack safety.

### 2020-09-09 rebuttal.txt

Sk2019b _define_ stack safety via a capability machine semantics that is supposed to embody stack
safety "by construction," and relating it to a standard semantics for the same machine via a full abstraction result.
We did not mean to imply that the same idea could not work for other machine models or enforcement mechanisms (although
this generalization is not actually in Sk2019b, even as suggested future work). Rather, we wished to distinguish our
approach from theirs by emphasizing that we define stack safety based not on any "correct by construction"
machine tied to some flavor of enforcement,  but rather on more abstract notions of observable behavior.

> * What exactly is the benefit of your approach to defining stack safety w.r.t. the approach of Skorstengaard 2019b?
>   Do you think it can be applied to more or less languages?
>   Do you think it is better suited or less well suited for reasoning about programs which depend on stack safety?
>   Do you think it is etc.
>
> * Do you think you could show safety of a specific example program that depends on stack safety (like the awkward example) and if so, how would such a proof work?
>
> * Do you think your approach to defining stack safety could be extended to languages with the following features (and if so, how?):
>   - varargs?  (seems problematic for call map)
SNA:

  Varargs can easily be implemented by establishing a more concrete calling convention in which
  the number of arguments for a call is apparent by examining the call state, at the cost of the
  convention-agnosticism of the call map. Alternatively the call map can take the entire state as
  an input and once again abstract over calling conventions. Once we can deduce the number of
  arguments of each call, our properties would work without alteration.

>   - dynamic code generation (seems problematic for the function map, call map and return map)
>   - dynamic linking (seems problematic for the function map, call map and return map)
>   - self-editing code (seems problematic for call map and return map)
>   - tail calls (seems potentially okay, but I'd like to be sure)
SNA:

  A tail call in terms of our properties would be a call for which confidentiality and integrity hold,
  but which is selectively exempted from Return Integrity. From the (tail) callees' perspective, it is
  permitted to return anywhere. But the return state for the callee is also that of its caller, and so
  the nested nature of the property means that if the caller obeys Return Integrity, then the callee
  returns to the correct place.

>   - exceptions and stack unwinding
>   - closures (seems potentially problematic for precision of the entry map)
>   How would you extend your approach to define stack safety in the presence of these features?
>
> * Do you think your approach implies the following properties?
>   - temporal safety in the presence of closures (i.e. closures should only be invokable by the adversary after the adversary first receives a reference to them)
>   - encapsulation of closures: it should not be possible for the adversary to invoke a closure with a different environment than the one they received?
>   If not, do you think it should imply these properties and if so, how do you think it could be extended?

SNA:
  Adding closures requires:
  - identifying code points where closures are created (i.e., adding another map)
  - adding machinery to track association of closure identities, entry points, and components
    containing valid closure pointers (via pointer provenance, as with both stack-derived and heao pointers)
  - extending WBCF to require that closure invocations respect the closure's designated entry point
    and use a valid closure pointer (ensuring temporal safety)
  - extending contours to treat some components as sealed, associated with closure identities
  - on closure invocation, set contour for encapsulated components to LC/LI, other components as normal for stack safety
    (together with above, ensuring encapsulation)
  - otherwise treat sealed components as HC/HI except that from a caller's perspective, they might change during
    nested calls (monotonicity is lost)

  The answers are yes, we believe that with addition of closures, natural extensions to WBCF and LSE would imply these
  properties. The addition of closures is very non-trivial but plausible.

### 2020-09-09 rebuttal.txt

Thank you for the thoughtful comments.  We first discuss the most important
points raised, then, in an optional appendix, respond point-by-point to each
individual piece of reviewer feedback.

@@

MOTIVATION

Reviewer C is particularly concerned with the way this work is
motivated. Our original goal in this work wasn't to formalize a stack safety
notion that is specific to micropolicies, but in hindsight we agree that the
result may appear to be tailored to that particular enforcement mechanism,
though we believe this is more a matter of presentation choices and
exposition-helping simplifications than a fundamental limitation. In
particular, we believe that the enforcement model can be used (with some
effort) to describe capability-based enforcement mechanisms, as we state in
Section 2.  However, given that this effort didn't make it into the final
product, we agree that the paper as it stands would be more accurately
described as an abstract property capturing micropolicy enforcement rather
than the all-encompassing stack safety notion we set out to state; we will
clarify this for a final version.

@@

RELATED WORK

All reviewers also wanted additional discussion of related work.  We agree,
in particular, that there should be a clearer comparison to the two
Skorstengaard et al papers, using the five informal criteria proposed
there. We plan to include a discussion like the following:

BCP: Can we format these so that it is clear which bits we're
getting (verbatim, even) from Skorstengaard and which are us?

BCP: Spell out abbreviations like LSE.  In general, we should not assume
reviewers remember all the details of the paper (much less the details of
Skorstengaard)!

(1) On intuition: The LSE part of the property we present aims to
characterize stack safety in security terms. For readers familiar with
integrity/confidentiality/noninterference, the eager version should be
fairly intuitive. The lazy/observable properties, though less intuitive, are
more interesting in that they precisely characterize behaviors that would be
considered violations to the more intuitive view of stack safety but that
can still be considered safe by an external observer. The WBCF part simply
formalizes standard behavior of stack-based control flow.

@@

By contrast, Skorstengaard relies on different prior intuitions:
for readers familiar with the behavior of pure stacks (or rather the
built-in oLCM stack), their property is more intuitive. For readers
versed in security, we believe there is value in a different
perspective.

(2) useful for reasoning: In order to reason about actual programs we
would indeed need to incorporate a "Meaning of Memory Safety"-style
frame rule. Our goal with this paper was to distill the essence of
stack safety without adding the complex (but solved) overhead of
dealing with heap locations. But the properties (and the testing
framework behind it) were designed with [Azevedo et al] in mind.

BCP: I'm not sure this is hitting the right point.  Clearly, to reason about
programs with both stack and heap, we need something like that frame rule.
But the reviewer might wonder about how our property can be used for
_reasoning_ about programs with just a stack...

(3) reusable in compiler chains: As you point out, this is not really
applicable, since this is a criterion tailored to what Skorstengaard
is aiming to do. We might consider what the equivalently tailored
criterion would be for us -- usable in "dynamic enforcement
environments" with potentially unsafe commodity compilers.
BCP: "you" is just one reviewer, but this is to be read by all three...

@@

We sketch how multiple threads/coroutines would work in the future
work section (l 1213-1219). In particular, we would need to further
distinguish stack locations as either "accessible" (current frame),
"belonging to the same call stack", or "belonging in another
stack". And then integrity would require that when a function changes
the locations in the "same" stack should be unchanged. Similarly for
cofnidentiality.

>  - In the future work section, the authors mentions several types of non-stack control flow that are not addressed by the current model, including tail calls. Other common non-stack control flow includes exceptions (such as c++ exceptions) and setjmp/longjmp. I understand that these non-stack control flow is more difficult to model, but for the exact reasons, how to formally characterize stack safety for these difficult cases becomes critical.

Tail calls were left out as it's only the characterization of the
return that poses a problem - once the point of context switch is
determined, the statement of stack safety remains largely the same.

BCP: Don't understand.  Are we saying tail calls are easy?  (Then why didn't
we just add them?)

@@

We propose two definitions of stack safety: one that is easier to understand
because it corresponds to a more naive view of what stack safety intuitively
means, and a slightly more complex and permissive one that (as Roessler and
Dehon showed) can be implemented more efficiently.  We believe both are of
interest.

> - This paper is lacking proper discussion on related work.
>
>    On the contrary to this paper's observation (L25), there're a few formalizations of stack safety (and rely/guarantee conditions for function invocation).
>       For example, see CompCert's assumption on function calls: https://github.com/AbsInt/CompCert/blob/94558ecb3e48261f12c644045d40c7d0784415e0/common/Events.v#L619

BCP: Obviously, this has to be addressed strongly and concretely.
LEO: I'm not really following the CompCert definition. Has anybody taking a crack at this?

There is some relation.  E.g., the ec_readonly condition

  https://github.com/AbsInt/CompCert/blob/94558ecb3e48261f12c644045d40c7d0784415e0/common/Events.v#L651

looks related.  But CompCert's definition (a) doesn't protect the stack at
all from _reading_, and (b) isn't "nested" the way ours is.  This makes
sense, since C doesn't protect the stack!

@@

> Another odd aspect of this paper is the fact that it doesn't propose one but three increasingly strong properties that characterize stack safety (see l86-96).
> What does this mean: why do we want stack safety to mean different things and which meaning is appropriate in which setting?

We view these stack safety properties as a spectrum. The main property
strives to capture an intuitive notion of stack safety, by relying on
abstract notions of observable behavior.  The stronger, lockstep
enforces a more conservative notion of stack safety, where almost any
digression from a canonical execution is considered unsafe. On the
other hand, the weaker lazy variant pushes the limit of the standard
intuition by allowing executions that would otherwise be unsafe, as
long as their effects are not actually observable. We think all
three are interesting in their own right and reveal interesting trade
offs in terms of enforcement or reasoning efficiency.

> Do the three variants score equally well on the above criteria?
> - For the "lockstep" variant, I get the impression that this is a proof-technical definition that is only used for proving the regularly stack-safety property for the micropolicy enforcement and not really important/useful in itself: is this also how you see it?

The lockstep variant is a slightly more technical artifact that is (1) easy
to enforce and (2) easy to test, but that is also interesting in its own
right because it captures a quite conservative notion of stack safety (while
still allowing the majority of reasonable stack-safe programs).

@@

See the full comparison in terms of the criteria in the summary.

> * Do you think you could show safety of a specific example program that depends on stack safety (like the awkward example) and if so, how would such a proof work?

LEO: This is the (4) criterion I guess again. I'm not sure exactly what this means.

BCP: I think this is a more focused question: Suppose you had a program that
you believed was correct according to this criterion and you wanted to prove
that it was.  How would the proof look?

@@

LEO: Polish needed here.
These three extensions would indeed be challenging. We could
potentially handle generated code by using special contours for the
instructions that generate it, which would correspond to a map that
allows some instructions to update the maps, including potentially
the map itself.

BCP: I don't actually believe the second sentence.

>   - tail calls (seems potentially okay, but I'd like to be sure)

Yes. The only minor problem with tail calls is characterizing the return;
once the point of context switch is determined, the statement of stack
safety remains largely the sam.

@@

Adding closures would require:
(1) identifying code points where closures are created
(2) adding machinery to track association of closure identities, entry points, and components
    containing valid closure pointers (including pointer provenance, as with stack-derived pointers)
(3) extending WBCF to require that closure invocations respect the closure's designated entry point
    and use a valid closure pointer (ensuring temporal safety and encapsulation)
(4) extending contours to treat some components as sealed, associated with closure identities
(5) on closure invocation, set contour for encapsulated components to LC/LI, other components as normal for stack safety
    (allowing access to closure)
(6) otherwise treat sealed components as HC/HI except that from a caller's perspective, they might change during
    nested calls (monotonicity is lost)

So to sum up, the answers are yes, we believe that to add closures,
natural extensions to WBCF and LSE would imply the properties
mentioned.

> * Are the lazy and strong variant of your properties specifically intended for the corresponding micropolicy-based enforcement techniques or do you believe they are more broadly applicable?

As discussed in the summary above and in your detailed comments, we did not
intend for the properties to be micropolicy specific, but rather to
represent a spectrum of stack safety properties presenting an interesting
trade-off in terms of enforcement or reasoning efficiency.

But we agree that the current exposition does not convey this generality,
and that the overall motivation of the work would be better portrayed as a
formal characterization of micropolicy-enforced stack safety.

### 2020-09-09 rebuttal.txt

>   - dynamic code generation (seems problematic for the function map, call map and return map)
>   - dynamic linking (seems problematic for the function map, call map and return map)
>   - self-editing code (seems problematic for call map and return map)

LEO: Polish needed here.
These three extensions would indeed be challenging. We could
potentially handle generated code by using special contours for the
instructions that generate it, which would correspond to a map that
allows some instructions to update the maps, including potentially
the map itself.

>   - tail calls (seems potentially okay, but I'd like to be sure)

Yes. The only minor problem with tail calls is characterizing the
return - once the point of context switch is determined, the statement
of stack safety remains largely the same. Notably a tail call's return
point is shared with its caller, so if the (non-tail-call) caller obeys
Return Integrity, the expected control flow of the tail call is already
captured.

@@

>   - closures (seems potentially problematic for precision of the entry map)

See below.

>   How would you extend your approach to define stack safety in the presence of these features?
>
> * Do you think your approach implies the following properties?
>   - temporal safety in the presence of closures (i.e. closures should only be invokable by the adversary after the adversary first receives a reference to them)
>   - encapsulation of closures: it should not be possible for the adversary to invoke a closure with a different environment than the one they received?
>   If not, do you think it should imply these properties and if so, how do you think it could be extended?

Adding closures would require: 
(1) identifying code points where closures are created
(2) adding machinery to track association of closure identities, entry points, and components
    containing valid closure pointers (including pointer provenance, as with stack-derived pointers)
(3) extending WBCF to require that closure invocations respect the closure's designated entry point
    and use a valid closure pointer (ensuring temporal safety)
(4) extending contours to treat some components as sealed, associated with closure identities
(5) on closure invocation, set contour for encapsulated components to LC/LI, other components as normal for stack safety
    (together with the above, ensuring encapsulation)
(6) otherwise treat sealed components as HC/HI except that from a caller's perspective, they might change during
    nested calls (monotonicity is lost)

### 2020-09-09 rebuttal.txt

Tail calls were left out as it's only the characterization of the
return that poses a problem - once the point of context switch is
determined, the statement of stack safety remains largely the same.

>  [1] Xiaoyang Xu, Masoud Ghaffarinia, Wenhao Wang, Kevin W. Hamlen, and Zhiqiang Lin. 2019. CONFIRM: evaluating compatibility and relevance of control-flow integrity protections for modern software. In Proceedings of the 28th USENIX Conference on Security Symposium (SEC’19). USENIX Association, USA, 18051821.
>
> Comments for author
> -------------------
> Please see the weaknesses section.
>
>
>
> Review #27B
> ===========================================================================
>
> Overall merit
> -------------
> D. Strong Reject
>
> Reviewer expertise
> ------------------
> Y. Knowledgeable
>
> Paper summary
> -------------
> This paper:
> - Defines stack safety as confidentiality, integrity, and well-bracketedness.
> - Evaluates existing stack protection mechanisms w.r.t. the defined notion of stack safety.
> - Proposes other protection mechanisms.
>
> Strengths
> ---------
> - This paper is clearly defining a notion of stack safety.
>
> Weaknesses
> ----------
> - The proposed notion of stack safety is too strong to be used for real-world programs.
> - The proposed notion of stack safety is too strong to be enforced by existing stack protection mechanisms.
> - This paper is lacking proper discussion on related work.
>
> Comments for author
> -------------------
> This paper is not clearly motivated, and as a result, has the following problems:
>
> - The proposed notion of stack safety is too strong to be used for real-world programs.
>
>   As discussed in Section 9, stack-derived pointers are not supported (L1199), but they are widely used in real-world programs.
>   Furthermore, without stack-derived pointers, stack safety is more-or-less obvious to define.

Leo: The sketch of stack-derived pointers in 1199 seems fairly intuitive to me. Double down on it?
APT: Does Sk2019b support stack-derived pointers? I don't think so, although I'm not certain.

@@

> - The proposed notion of stack safety is too strong to be enforced by existing stack protection mechanisms.
>
>   [Roessler and DeHon 2018]'s "lazy tagging and clearing" is itself not stack-safe according to this paper.
>   To understand the mechanism, this paper additionally defines the "observable" versions of stack safety.
>   But then why you didn't use the observable version as the authoritative definition?

We propose two definitions of stack safety: one that is easier to understand
because it corresponds to a more naive view of what stack safety intuitively
means, and a slightly more complex and permissive one that (as Roessler and
Dehon showed) can be implemented more efficiently.  We believe both are of
interest.

> - This paper is lacking proper discussion on related work.
>
>    On the contrary to this paper's observation (L25), there're a few formalizations of stack safety (and rely/guarantee conditions for function invocation).
>       For example, see CompCert's assumption on function calls: https://github.com/AbsInt/CompCert/blob/94558ecb3e48261f12c644045d40c7d0784415e0/common/Events.v#L619

BCP: Obviously, this has to be addressed strongly and concretely.
LEO: I'm not really following the CompCert definition. Has anybody taking a crack at this?
APT: The CompCert specification cited here characterizes the assumptions that the comiler
correctness proof makes about the behavior of external functions.  
When taken together with other parts of the CompCert spec that
describe the block memory model and the treatment of control stacks,
this does imply some facts about the stack safety conditions that CompCert
assumes are provided by the external environment. But it is certainly not intended
as a general characterization of stack safety, and it does not seem suitable
as a basis for checking validity of an enforcement mechanism.

### 2020-09-10 rebuttal.txt

> * What exactly is the benefit of your approach to defining stack safety w.r.t. the approach of Skorstengaard 2019b?
>   Do you think it can be applied to more or less languages?
>   Do you think it is better suited or less well suited for reasoning about programs which depend on stack safety?
>   Do you think it is etc.

See the full comparison in terms of the criteria in the summary.

> * Do you think you could show safety of a specific example program that depends on stack safety (like the awkward example) and if so, how would such a proof work?

LEO: This is the (4) criterion I guess again. I'm not sure exactly what this means.
APT: I agree with SNA that this is the (2) criterion. Maybe we could doctor up the example program to
include some external call to an adversary and sketch how it we would use it to prove 
that it still prints 0, 60 (or whatever).

@@

LEO: Polish needed here.
These three extensions would indeed be challenging. We could
potentially handle generated code by using special contours for the
instructions that generate it, which would correspond to a map that
allows some instructions to update the maps, including potentially
the map itself.

BCP: I don't actually believe the second sentence.

>   - tail calls (seems potentially okay, but I'd like to be sure)

Yes. The only (minor) problem with tail calls is characterizing the
return - once the point of context switch is determined, the statement
of stack safety remains largely the same. Notably a tail call's return
point is shared with its caller, so if the (non-tail-call) caller obeys
Return Integrity, the expected control flow of the tail call is already
captured.
